---
title: "Problem Set 4"
author: "Alexandra Dobbins"
date: "2/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(gt)
library(dplyr)
library(fivethirtyeight)
library(ggplot2)

#Eliminates warning messages about varaibles being character/numeral, etc. 
nc09 <- read_csv(file = "ps_4_elections-poll-nc09-3_copy.csv", 
                  col_types =  cols(
                   .default = col_character(),
                   turnout_scale = col_double(),
                   turnout_score = col_double(),
                   w_LV = col_double(),
                   w_RV = col_double(),
                   final_weight = col_double(),
                   timestamp = col_datetime(format = "")))

```


```{r question 1}

#Question 1

#Initial cleaning of data to use for future calculations
nc_09 <-
  nc09 %>% 
  #isoalte variables of interest for question 1
    select(response, race_eth, final_weight) %>%
  #group by reponse and race to make data neat
    group_by(response, race_eth) %>% 
  #calcualte sum of all the weights of voters 
    summarize(total = sum(final_weight)) %>% 
  #remove non-responsive data points from ethnicity varaible
    filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%
  #spread by response to create columns for each response category
    spread(key = response, value = total) %>% 
  #one of the Asian responses contained an NA, so change it to 0 so data remains in set
    replace_na(list(Und = 0)) %>% 
  #create varaible all and recrate Dem, Rep, and Und to represent percentage of only those three categories 
    mutate(all = Dem + Rep + Und) %>% 
      mutate(Dem = Dem/all) %>% 
      mutate(Rep = Rep/all) %>% 
      mutate(Und = Und/all) %>% 
    select(race_eth, Dem, Rep, Und) %>% 
  #ungroup so mutations can be done
    ungroup() %>% 
  #manually choose order of race varaibles for chart
    mutate(race_eth = fct_relevel(race_eth, "White", "Black", "Hispanic", "Asian", "Other")) %>% 
    arrange(race_eth) 

#Mad Libs


Dem_Support <- 
  #Count number of people who reponded Democrat, print only the numeric value in n column
  nc09 %>% 
  filter(response == "Dem") %>% 
  nrow()

RepCount <- nc09 %>% 
  #Count number of people who reponded Republican, print only the numeric value in n column
  filter(response == "Rep") %>% 
  nrow()

UndCount <- nc09 %>% 
  #Count number of people who reponded Und, print only the numeric value in n column
  filter(response == "Und") %>% 
  nrow()

GenderDiff <- 
  nc09 %>%
  #Select the two columns of interest
  select(gender, gender_combined) %>% 
  #filter for columns where the inputs are unequal
  filter(gender != gender_combined) %>% 
  #print numeric count value of n column
  nrow()

White <-
  nc09 %>% 
  #Select the two columns of interest
  select(race_eth, file_race_black) %>% 
  #filter for columns where the input is race_eth is With, then columns where inputs are unequal
  filter(race_eth == "White", race_eth != file_race_black) %>%
  #print numeric count value of n column
  nrow()
  
#Create two variables (to be subtracted) which select for the first Dem timestamp, and first Rep timestamp
dem_time <- nc09 %>% filter(response == "Dem") %>%  slice(1) %>%  pull(timestamp)
rep_time <- nc09 %>% filter(response == "Rep") %>%  slice(1) %>%  pull(timestamp)

```

**Question 1:**

There were `r Dem_Support` respondents who supported the Democratic candidate.

There were `r RepCount - UndCount` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables `gender` and `gender_combined`. There are `r GenderDiff` individuals for whom these variables have different values.

There are `r White` respondents listed as "White" under `race_eth` who are not listed as "White" under `file_race_black`.

The first response of Rep came `r round(rep_time - dem_time)` minutes (rounded to the nearest minute) before the first response of Dem.



**Question 2:**

```{r}
#Question 2

nc_09 %>% 
  #use gt() to create a chart
  gt() %>% 
  #name chart
  tab_header(title = "North Carolina's 9th Congressional District Voting Distribution by Race and Party")%>% 
  #rename column variables to match target chart
  cols_label(
    #I think there needed to be a label on the ethnicity column 
    race_eth = "Ethnicity",
    Dem = "DEM.",
    Rep = "REP.",
    Und = "UND."
    ) %>%
  #transform data points to percentages with no decimal points
  fmt_percent(columns = vars(Dem, Rep, Und), decimals = 0) %>% 
  #replace NA with a dash for Asian x Und
  na_if(0) %>% 
  fmt_missing(columns = vars(Und), rows = 4)

```



**Question 3:**

```{r}
#Question 3

nc_educ <-
  nc09 %>% 
  #select for relevant variables
    select(educ, final_weight) %>%
  #remove people who did not respond to educ column
    filter(educ != "[DO NOT READ] Refused") %>%
  #manually choose order of varaibles on graph
    mutate(educ = fct_relevel(educ, "Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))

nc_educ %>% 
  #create a ggplot of voter weight by education 
  ggplot(aes(x = educ, y = final_weight)) +
  #violin plot to show densities of count
  geom_violin()+
  #flip axis so labels can be read/match target plot
  coord_flip()+
  #add in dots to represent actual data points, with width, shade, and size adjusted to match target plot
  geom_jitter(aes(x = educ, y = final_weight), width = .24, size = 1, alpha = .5)+
  #name plot
  labs(title = "More Educated Matter Less in North Carolina 9th", subtitle = "Poll gives more weight to people who are less likely to participate in polls", caption = "New York Times Upshot/Siena College 2018 live polls")+
  #name x and y variables
  ylab("Weight Given to Respondent in Calculating Poll Results")+
  xlab("")

```

Question 4:

```{r }

#Question 4

q4_graph <-
  nc09 %>% 
  #select for relevant varaibles to look at how education effects people's intentions of voting
  select(educ, likely) %>% 
  #filter out repondants in educ and likely categories who did not respond 
  filter(educ != "[DO NOT READ] Refused", likely != "[DO NOT READ] Don't know/Refused") %>%
  #manually choose order of varaibles for fill 
  mutate(educ = fct_relevel(educ, "Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree")) %>% 
  #manually choose order of varaibles on graph
  mutate(likely = fct_relevel(likely, "Not at all likely", "Not very likely","Somewhat likely", "Very likely", "Almost certain", "Already voted"))


q4_graph %>% 
  ggplot() +
  #create bar plot with likihood of voting on x axis and colored to represent education level. alpha lowered to .5 to make bars a little transparent and pretty
  geom_bar(mapping = aes(x = likely, fill = educ), alpha = .7)+
  #flip so labels can be read
  coord_flip()+
  #rename axis
  ylab("Citizens (number)")+
  xlab("Self-Reported Odds of Voting")+
  #name chart
  labs(title = "Self-Reported Odds of Voter Participation with Respect to Education", 
       subtitle = "An exploration of education's effects on political participation")+
  #write directly on graph to point out interesting response which Kieran Healy would like very much 
  annotate(geom = "text", x = .85, y = 90, size = 2.5, label = "<- Surprisingly honest feedback about voting intentions")

```
